Ionization of molecules by multicharged bared ions by using the stoichiometric model

### 2019-09-13 NewMonti25.tex
version25 sent
scaling figure + small arranges in text

Lets us consider a molecule $M$ composed by $n_{\alpha}$ atoms of the
element $\alpha$, the SSM describes the total ionization cross section 
of the molecule $\sigma_{M}$ as a simple sum of ionization cross 
sections of the isolated atoms $\sigma_{\alpha}$, 
\begin{equation}
 \sigma_{M}=\sum\limits_{\alpha}n_{\alpha}\sigma_{\alpha}\,.  
 \label{eq:sumion}
\end{equation}
We divided sixteen molecular targets of our interest in three families. 
The classification defined is given in Table~\ref{tab:families}.
\begin{table}
\begin{center}
\begin{tabular}{|p{0.06\textwidth}|p{0.65\textwidth}|}
\hline
 CH  & CH$_4$, C$_2$H$_2$, C$_2$H$_4$, C$_2$H$_6$, and C$_6$H$_6$ \\
\hline
 CHN & C$_5$H$_5$N, C$_4$H$_4$N$_2$, C$_2$H$_7$N, and CH$_5$N \\
\hline
 \multirow{4}{*}{DNA} & adenine (C$_5$H$_5$N$_5$), 
                        cytosine (C$_4$H$_5$N$_2$O), \\
     & guanine (C$_5$H$_5$N$_5$O), thymine (C$_5$H$_6$N$_2$O$_2$), \\
     & uracil (C$_4$H$_4$N$_2$O$_2$), DNA backbone (C$_5$H$_{10}$O$_5$P), \\
     & dry DNA (C$_{20}$H$_{27}$N$_7$O$_{13}$P$_2$)\\
\hline
\end{tabular}
\caption{Sixteen molecular targets of our interest classified in three 
families.}
\label{tab:families}
\end{center}
\end{table}

@@

The first attempt to develop a comprehensive but straightforward 
phenomenological model for electron ejection from large molecules was 
proposed by Toburen {\it et al.}~\cite{toburen1975,toburen1976}. 
The authors found it convenient to scale the experimental ionization 
cross section in terms of the number of outer or weakly bound valence 
electrons (total number of electrons minus the inner-shell ones). 
We introduce the cross section per weakly bound electron or ionization 
cross section per valence electron, $\sigma_{e}$, as
\begin{equation}
\sigma_{e}=\frac{\sigma_{M}}{N_{M}}=\frac{\sum\limits_{\alpha}
n_{\alpha}\sigma_{\alpha}}{\sum\limits_{\alpha}n_{\alpha}\nu_{\alpha}}
=\sigma_{e}(v)\,, 
\label{27} 
\end{equation}
where
\begin{equation}
\nu_{\alpha}=\left\{ 
\begin{array}{ll}
4, & \text{for C,} \\ 
5, & \text{for N and P,} \\ 
6, & \text{for O and S,}
\end{array}\right.
\label{eq:nelec} 
\end{equation}
and $\nu_{\text{H}}=1$. The Toburen rule can be stated by saying that 
$\sigma_{e}$ is a \textit{universal} parameter independent on the 
molecule, which depends solely on the impact velocity. 
In Ref.~\cite{toburen1976}, it was found experimentally that for proton 
impact on some simple molecules, $\sigma_{e}$ results
\begin{equation}
\begin{array}{cccc}
E\text{(MeV)}               & 0.25 & 1.00 & 2.00 \\ 
\sigma_{e} (10^{-16}cm^{2}) & 0.39 & 0.17 & 0.11
\end{array}
\label{30}
\end{equation}
with an estimated error of about 20\%. A similar ratio was found in 
Ref.~\cite{itoh2013} for proton impact on uracil and adenine. 
Figure 4a shows $\sigma_{e}$ calculated with the CDW as a function of 
the impact velocity for different projectile charges computed with the 
SSM for the sixteen molecular targets displayed in 
Table~\ref{tab:families}. The universality with $Z$ is the one provided 
by Born approximation, i.e., $\sigma _{e}(Z)=Z^{2}\sigma_{e}(Z=1)$, 
and it holds for large impact velocities, as shown Figure 4a.
Of course, for lower impact velocities, the CDW breaks the behavior of 
the $Z^{2}$ rule. Although the Toburen rule holds for high energies, 
its performance is still not satisfactory: the universal band is quite 
broad. The departure of our theoretical 
results from the Toburen rule can be easily explained by the fact that 
our atomic cross sections $\sigma_{\alpha}$ behave differently for 
each atom in terms of the projectile charges. By inspecting Figure 1, 
one can easily see that the rule $\sigma_{\alpha}/\nu_{\alpha}\sim
\sigma_{e}$ is not well satisfied by the CDW. A much better general 
atomic rule is given by
\begin{equation}
\sigma_{e}^{\text{CDW}}=\frac{\sum\limits_{\alpha}
n_{\alpha}\sigma_{\alpha}^{\text{CDW}}}{\sum\limits_{\alpha}n_{\alpha}
\nu_{\alpha}^{\text{CDW}}},
\label{32} 
\end{equation}
where 
\begin{equation}
\nu_{\alpha }^{\text{CDW}} \sim\left\{ 
\begin{array}{ll}
4,\ \ \  & \text{for C, N, and O,} \\ 
4.5,\ \  & \text{for P and S}
\end{array}
\right. 
\label{eq:scalingCDW}
\end{equation}
and, obviously, $\nu_{\text{H}}^{\text{CDW}}=1$. The cross sections 
$\sigma_{e}^{\text{CDW}}$ are plotted in Figure~4b. A much better sharp 
band is observed, especially at high energy, where the theory is 
expected to work. It will be interesting for experiments to cross-check 
this theoretical prediction.

@@

\begin{figure}[H]
\centering
\includegraphics[width=0.95\textwidth]{energylevels/DNA.eps}
\end{figure}

@@

We calculated ionization cross sections by impact of antiprotons, 
H$^{+}$, He$^{2+}$, Be$^{4+}$, C$^{6+}$, and O$^{8+}$ for sixteen 
molecules containing H, C, N, O, P and S with the CDW method. The 
importance of the influence of $Z$ was observed in the mean energy 
$\overline{E}_{\alpha}$ and angle $\overline{\theta}_{\alpha}$. 
For a given target $\alpha$, as $Z$ increases, $\overline{E}_{\alpha}$ 
increases, but $\overline{\theta}_{\alpha}$ decreases. At high impact 
energy, say larger than 1 MeV/amu, these values converge to the Born 
approximation, which embodies the simple $Z^{2}$ law. The sixteen 
molecules selected were investigated using the simple stoichiometric 
model. Results for six DNA bases were presented and compared with the 
sparse available experiments. We explore the rule of Toburen which scales 
all the molecular ionization cross section when divided by the number of 
weakly bound valence electrons $\nu_{\alpha}$ given by Eq.~(\ref{27}). 
We found the rule scales much better when normalizing our theoretical 
ionization cross sections to the number $\nu_{\alpha}^{\text{CDW}}$ 
given by Eq.~(\ref{eq:scalingCDW}). Finally, we attempt to improve the 
stoichiometric model by using the Mulliken charge to define a new model 
containing fractional rather than integer proportions. No substantial 
correction was found, which indicates that the SSM works quite well.
By inspecting the molecular binding energy from quantum mechanical 
structure calculations, we were able to understand the values defined
for $\nu_{\alpha}^{\text{CDW}}$.

>>>>

Lets us consider a molecule $M$ composed by $n_{\alpha}$ atoms of the
element $\alpha$, the SSM describes the total ionization cross section 
of the molecule $\sigma_{M}$ as a simple sum of ionization cross 
sections of the isolated atoms $\sigma_{\alpha}$, 
\begin{equation}
 \sigma_{M}=\sum\limits_{\alpha}n_{\alpha}\sigma_{\alpha}\,.  
 \label{eq:sumion}
\end{equation}
We divided sixteen molecular targets of our interest in three families. 
The classification defined is given in Table~\ref{tab:families}.
\begin{table}
\begin{center}
\begin{tabular}{|p{0.06\textwidth}|p{0.62\textwidth}|}
\hline
 CH  & CH$_4$, C$_2$H$_2$, C$_2$H$_4$, C$_2$H$_6$, and C$_6$H$_6$ \\
\hline
 CHN & C$_5$H$_5$N, C$_4$H$_4$N$_2$, C$_2$H$_7$N, and CH$_5$N \\
\hline
 \multirow{4}{*}{DNA} & adenine (C$_5$H$_5$N$_5$), 
                        cytosine (C$_4$H$_5$N$_2$O), \\
     & guanine (C$_5$H$_5$N$_5$O), thymine (C$_5$H$_6$N$_2$O$_2$), \\
     & uracil (C$_4$H$_4$N$_2$O$_2$), tetrahydrofuran (C$_4$H$_8$O), \\
     & DNA backbone (C$_5$H$_{10}$O$_5$P), \\
     & dry DNA (C$_{20}$H$_{27}$N$_7$O$_{13}$P$_2$)\\
\hline
\end{tabular}
\caption{Seventeen molecular targets of our interest classified in three 
families.}
\label{tab:families}
\end{center}
\end{table}

@@

The first attempt to develop a comprehensive but straightforward 
phenomenological model for electron ejection from large molecules was 
proposed by Toburen {\it et al.}~\cite{toburen1975,toburen1976}. 
The authors found it convenient to scale the experimental ionization 
cross section in terms of the number of outer or weakly bound valence 
electrons (total number of electrons minus the inner-shell ones). 
We introduce the cross section per weakly bound electron or ionization 
cross section per valence electron, $\sigma_{e}$, as
\begin{equation}
\sigma_{e}=\frac{\sigma_{M}}{N_{M}}=\frac{\sum\limits_{\alpha}
n_{\alpha}\sigma_{\alpha}}{\sum\limits_{\alpha}n_{\alpha}\nu_{\alpha}}
=\sigma_{e}(v)\,, 
\label{27} 
\end{equation}
where $\nu_{\alpha}$ is the number of weakly bound electrons of the atom
$\alpha$, given by
\begin{equation}
\nu_{\alpha}=\left\{ 
\begin{array}{ll}
4, & \text{for C,} \\ 
5, & \text{for N and P,} \\ 
6, & \text{for O and S,}
\end{array}\right.
\label{eq:nelec} 
\end{equation}
and $\nu_{\text{H}}=1$. The Toburen rule can be stated by saying that 
$\sigma_{e}$ is a \textit{universal} parameter independent on the 
molecule, which depends solely on the impact velocity. 
In Ref.~\cite{toburen1976}, it was found experimentally that for proton 
impact on some simple molecules, $\sigma_{e}$ results
\begin{equation}
\begin{array}{cccc}
E\text{(MeV)}               & 0.25 & 1.00 & 2.00 \\ 
\sigma_{e} (10^{-16}cm^{2}) & 0.39 & 0.17 & 0.11
\end{array}
\label{30}
\end{equation}
with an estimated error of about 20\%. A similar ratio was found in 
Ref.~\cite{itoh2013} for proton impact on uracil and adenine.

\begin{figure}[t!]
\includegraphics[width=\textwidth]{figuras/scaling85.eps}
\caption{Ionization cross section per weakly bound electron using
(a) Eq.~(\ref{eq:nelec}) and (b) Eq.~(\ref{eq:scalingCDW}).}
\label{fig:newscaling}
\end{figure}
Figure 4a shows $\sigma_{e}$ calculated with the CDW as a function of 
the impact energy for different projectile charges computed with the 
SSM for the seventeen molecular targets displayed in 
Table~\ref{tab:families}. The universality with $Z$ is the one provided 
by Born approximation, i.e., $\sigma _{e}(Z)=Z^{2}\sigma_{e}(Z=1)$, 
and it holds for large impact velocities, as shown Figure 4a.
Of course, for lower impact velocities, the CDW breaks the behavior of 
the $Z^{2}$ rule. Although the Toburen rule holds for high energies, 
its performance is still not satisfactory: the universal band is quite 
broad. The departure of our theoretical 
results from the Toburen rule can be easily explained by the fact that 
our atomic cross sections $\sigma_{\alpha}$ behave differently for 
each atom in terms of the projectile charges. By inspecting Figure 1, 
one can easily see that the rule $\sigma_{\alpha}/\nu_{\alpha}\sim
\sigma_{e}$ is not well satisfied by the CDW. A much better general 
atomic rule is given by
\begin{equation}
\sigma_{e}^{\text{CDW}}=\frac{\sum\limits_{\alpha}
n_{\alpha}\sigma_{\alpha}^{\text{CDW}}}{\sum\limits_{\alpha}n_{\alpha}
\nu_{\alpha}^{\text{CDW}}},
\label{32} 
\end{equation}
where 
\begin{equation}
\nu_{\alpha }^{\text{CDW}} \sim\left\{ 
\begin{array}{ll}
4,\ \ \  & \text{for C, N, and O,} \\ 
4.5,\ \  & \text{for P and S}
\end{array}
\right. 
\label{eq:scalingCDW}
\end{equation}
and, obviously, $\nu_{\text{H}}^{\text{CDW}}=1$. The cross sections 
$\sigma_{e}^{\text{CDW}}$ are plotted in Figure~4b. A much better sharp 
band is observed, especially at high energy, where the theory is 
expected to work. It will be interesting for experiments to cross-check 
this theoretical prediction.

@@

\begin{figure}[H]
\centering
%\includegraphics[width=0.95\textwidth]{energylevels/DNA.eps}
\includegraphics[width=0.95\textwidth]{DNA.eps}
\end{figure}

@@

We calculated ionization cross sections by impact of antiprotons, 
H$^{+}$, He$^{2+}$, Be$^{4+}$, C$^{6+}$, and O$^{8+}$ for seventeen 
molecules containing H, C, N, O, P and S with the CDW method. The 
importance of the influence of $Z$ was observed in the mean energy 
$\overline{E}_{\alpha}$ and angle $\overline{\theta}_{\alpha}$. 
For a given target $\alpha$, as $Z$ increases, $\overline{E}_{\alpha}$ 
increases, but $\overline{\theta}_{\alpha}$ decreases. At high impact 
energy, say larger than 1 MeV/amu, these values converge to the Born 
approximation, which embodies the simple $Z^{2}$ law. The seventeen 
molecules selected were investigated using the simple stoichiometric 
model. Results for six DNA bases were presented and compared with the 
sparse available experiments. We explore the rule of Toburen which scales 
all the molecular ionization cross section when divided by the number of 
weakly bound valence electrons $\nu_{\alpha}$ given by Eq.~(\ref{27}). 
We found the rule scales much better when normalizing our theoretical 
ionization cross sections to the number $\nu_{\alpha}^{\text{CDW}}$ 
given by Eq.~(\ref{eq:scalingCDW}). Finally, we attempt to improve the 
stoichiometric model by using the Mulliken charge to define a new model 
containing fractional rather than integer proportions. No substantial 
correction was found, which indicates that the SSM works quite well.
By inspecting the molecular binding energy from quantum mechanical 
structure calculations, we were able to understand the values defined
for $\nu_{\alpha}^{\text{CDW}}$.
